Variants on the Flight and Maintenance Planning problem

### 2018-07-19 latex/notes/model/model.tex
added constraints of availability to model formulation in latex

\usepackage{epsfig}
%\usepackage{timesnew}
% \usepackage{harvard}
\usepackage[]{natbib}
\usepackage{amsmath}
\usepackage{booktabs}

\usepackage[latin1]{inputenc}
\usepackage{aeguill}
\usepackage[frenchb,english]{babel}
\usepackage{caption}

\setlength{\parskip}{2ex} \pagestyle{myheadings}
\markright{\hspace*{3.8cm} \textit{MOSIM18 - June 27-29, 2018 - Toulouse - France}}

@@

    \begin{tabular}{p{5mm}p{170mm}}
        % $\mathcal{I}$     &  resources. \\
        % $\mathcal{T}$     &  time periods inside planning horizon. \\
        % $\mathcal{J}$     &  tasks. \\
        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$. \\
        % $\mathcal{T}^{st}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - S+1\}},  ..., {t}$\}. \\
        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - Ret^{m}+1\}},  ..., {t}\}$. \\
        $\mathcal{T}^{M}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - Ret^{M}+1\}},  ..., \max{\{1, t - Ret^{m}+1\}}$\}. \\
        $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$. \\
    \end{tabular}

@@

    \begin{tabular}{p{8mm}p{167mm}}
        $H_j$             & amount of resource time required by task $j$. \\
        $R_j$             & number of resources required by task $j$. \\
        $MT_j$            & minimum number of consecutive periods a resource can be assigned to task $j$. \\
        % $aut_i$           & maximal absolute usage time for resource $i$. \\
        % $aet_i$           & maximal absolute elapsed time for resource $i$. \\
        $M$               & maintenance duration in number of periods. \\
        $M^{max}$         & maximum number of parallel maintenance operations. \\
        $E$               & remaining elapsed time after a maintenance. \\
        $H$               & remaining usage time after a maintenance. \\
        % $S$               & Storage duration in number of periods. \\
        $W_1$             & weight of the first objective in the objective function. \\
        $W_2$             & weight of the second objective in the objective function. \\
        $N_t$             & number of resources in already-planned maintenances in period $t$ at the beginning of the planning horizon.\\
        $D_t$             & number of resources to be assigned in total in period $t$. \\
        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. \\
        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
        % $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
        % $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
        $Ret^{M}$         & maximum number of periods between two consecutive maintenances. \\
        $Ret^{m}$         & minimum number of periods between two consecutive maintenances. \\
    \end{tabular}

@@

    The objective is to simultaneously minimize the maximum number of maintenances and the maximum number of unavailable aircraft.

@@

    where we want to minimize the total number of resource hours given via maintenances and we want to maximize the final state measured in available hours of resources. These two objectives have the same units and can be easily compared and they ensure the resource hours are used in the most efficient way possible.
    % where weights $W_1$ and $W_2$ are chosen by the decision maker. $W_1$ penalizes the maximum number of aircraft in maintenance among all periods. $W_2$ penalizes the maximum number of unavailable aircraft among all periods.
    The following constraints are used in the model:       
    \begin{align}
        % maximum capacity1:
        % & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
        %   & t \in \mathcal{T} \label{eq:capacity1}\\
        & \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \leq M^{max} \label{eq:capacity1}\\
        % availability
       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
        &t \in \mathcal{T} \label{eq:avalaibility1}\\
        % min assignments:
        & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
                & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
        % just doing one thing at any given time:
        & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
    \end{align}

    Maintenance capacity is controlled by (\ref{eq:capacity1}). The number of unavailable resources is defined by (\ref{eq:avalaibility1}). Tasks' resource requirements are defined by (\ref{eq:taskres}). Constraints (\ref{eq:state}) guarantee that a resource can be used only for one task, a maintenance operation or a storage operation at the same period.

@@

    Constraints (\ref{eq:start1}) controls the definition of starting a consecutive assignment. If a task $j$ is assigned in period $t$ and was not assigned the previous period ($t-1$).
    Constraints (\ref{eq:start2}) controls the minimum duration of a consecutive task assginment. If an assignment to task $j$ is started in period $t$, it needs to continue in periods $t' \in \mathcal{T}^{MT}_{jt}$.

@@

        % These constraints calculate the balances of hours for each resource.
    The remaining usage time is defined by (\ref{eq:rut_upper})-(\ref{eq:rut_initial}) and its limits by (\ref{eq:rut_lower})-(\ref{eq:mu}). 
    The remaining elapsed time is defined by (\ref{eq:ret_min})-(\ref{eq:ret_max}).

>>>>

\usepackage{epsfig}
%\usepackage{timesnew}
% \usepackage{harvard}
% \usepackage[]{natbib}
\usepackage{amsmath}
\usepackage{booktabs}

\usepackage[latin1]{inputenc}
% \usepackage{aeguill}
\usepackage[frenchb,english]{babel}
\usepackage{caption}

\setlength{\parskip}{2ex} \pagestyle{myheadings}
% \markright{\hspace*{3.8cm} \textit{MOSIM18 - June 27-29, 2018 - Toulouse - France}}

@@

    \begin{tabular}{p{5mm}p{170mm}}
        $\mathcal{I}$     &  resources. \\
        $\mathcal{T}$     &  time periods inside planning horizon. \\
        $\mathcal{J}$     &  tasks. \\
        $\mathcal{K}$     &  cluster of resources that share the same capacities to do tasks. \\
        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
        $\mathcal{I}_k$     &  resources $i \in \mathcal{I}$ that are included in cluster $k$. \\
        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$. \\
        % $\mathcal{T}^{st}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - S+1\}},  ..., {t}$\}. \\
        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - Ret^{m}+1\}},  ..., {t}\}$. \\
        $\mathcal{T}^{M}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - Ret^{M}+1\}},  ..., \max{\{1, t - Ret^{m}+1\}}$\}. \\
        $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$. \\
    \end{tabular}

@@

    \begin{tabular}{p{8mm}p{167mm}}
        $H_j$             & amount of resource time required by task $j$. \\
        $R_j$             & number of resources required by task $j$. \\
        $MT_j$            & minimum number of consecutive periods a resource can be assigned to task $j$. \\
        % $aut_i$           & maximal absolute usage time for resource $i$. \\
        % $aet_i$           & maximal absolute elapsed time for resource $i$. \\
        $M$               & maintenance duration in number of periods. \\
        $M^{max}$         & maximum number of parallel maintenance operations. \\
        $E$               & remaining elapsed time after a maintenance. \\
        $H$               & remaining usage time after a maintenance. \\
        % $S$               & Storage duration in number of periods. \\
        % $W_1$             & weight of the first objective in the objective function. \\
        % $W_2$             & weight of the second objective in the objective function. \\
        % $N_t$             & number of resources in already-planned maintenances in period $t$ at the beginning of the planning horizon.\\
        % $D_t$             & number of resources to be assigned in total in period $t$. \\
        $AK_{kt}$         & maximum number of resources in cluster $k$ that can to be under maintenance at period $t$.\\
        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. \\
        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
        % $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
        % $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
        $Ret^{M}$         & maximum number of periods between two consecutive maintenances. \\
        $Ret^{m}$         & minimum number of periods between two consecutive maintenances. \\
    \end{tabular}

@@

    The objective is to minimize the number of maintenances and maximize the end status of the resources.

@@

    where we want to minimize the total number of resource hours given via maintenances and we want to maximize the final state measured in available hours of resources. These two objectives have the same units and can be easily compared and they ensure the resource hours are used in the most efficient way possible.
    % where weights $W_1$ and $W_2$ are chosen by the decision maker. $W_1$ penalizes the maximum number of resources in maintenance among all periods. $W_2$ penalizes the maximum number of unavailable resources among all periods.
    The following constraints are used in the model:       
    \begin{align}
        % maximum capacity1:
        % & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
        %   & t \in \mathcal{T} \label{eq:capacity1}\\
        & \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \leq M^{max} \label{eq:capacity1}\\
        % availability
       % & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
       %  &t \in \mathcal{T} \label{eq:avalaibility1}\\
        % min assignments:
        & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
                & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
        % just doing one thing at any given time:
        & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
    \end{align}

    Maintenance capacity is controlled by (\ref{eq:capacity1}). Tasks' resource requirements are defined by (\ref{eq:taskres}). Constraints (\ref{eq:state}) guarantee that a resource can be used only for one task, a maintenance operation or a storage operation at the same period.

@@

    Constraints (\ref{eq:start1}) controls the definition of starting a consecutive assignment. If a task $j$ is assigned in period $t$ and was not assigned the previous period ($t-1$).
    Constraints (\ref{eq:start2}) controls the minimum duration of a consecutive task assginment. If an assignment to task $j$ is started in period $t$, it needs to continue in the following periods: $t' \in \mathcal{T}^{MT}_{jt}$.

    \begin{align}
       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} \leq AK_{kt}
        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-cluster}
    \end{align}

    Constraints (\ref{eq:avalaibility-cluster}) control a minimum availability for each cluster $k$ of resources. A cluster is defined by the most general group of resources that is required, exclusively, by at least one task.

@@

        % These constraints calculate the balances of hours for each resource.
    The remaining usage time is defined by (\ref{eq:rut_upper})-(\ref{eq:rut_initial}) and its limits by (\ref{eq:rut_lower})-(\ref{eq:mu}).

    The minimum elapsed time is defined by (\ref{eq:ret_min}). If a maintenance is done in some period $t$, we know that it cannot be done in the periods $t'$ that are close to it.

    On the other hand, due to the maximum elapsed time between maintenances, constraints (\ref{eq:ret_max}) force that, if we have a maintenance on period $t$, we have to have done at least another in one far away period $t'$.
